On the use of petri nets for modelling high-level programs

### 2010-07-20 Documents/Publications/type-2/type-2.tex
type-2: using firing sequences when proving liveness and safety

  Despite the restrictions imposed by the production rules, the Type-2
  class of petri nets is powerful enough to represent programs written
  in traditional imperative languages such as C/C++ and also in
  synchronous languages such as Esterel. It has been used as a
  transition step in translating imperative programs to hardware
  descriptions\cite{ahir-thesis}. The resulting high-level synthesis
  flow uses the labelling scheme to identify opportunities for sharing
  hardware resources in the absence of timing information.

@@

@@

Consider a simple place $p$ in a live and safe petri net $N =
(P,T,E,M)$. Let $u,v$ be the predecessor and successor transitions of
$p$ in $N$, i.e, $u,v \in T$ and $(u,p), (p,v)\in E$. Since $N$ is
live and safe, it follows that when $u$ fires, $v$ must fire before
$u$ may fire again (safety), and $v$ must eventually fire when $u$
fires (liveness). Thus, any firing sequence of the petri net $N$ that
includes the firing of $u$ (and consequently also $v$) effectively
puts a token in $p$ and then eventually removes it. Thus, the petri
net $N' = N - \{p,u,v\}$ can be modelled as an initially marked simple
place $p_0$ that forms a cycle with $u$, $v$ and $p$. Similarly,
relative to a simple transition $t$, the rest of the petri net can be
modelled as an initially marked simple place $p_0$ that forms a simple
cycle with $t$. Both replacements are valid if and only if the
original petri net is live and safe; the resulting simple cycle is
also live and safe. We use this simplification to examine the effect
of applying a construction rule on a Type-1 petri net:

>>>>

  Despite the restrictions imposed by the production rules, the Type-2
  class of petri nets is powerful enough to represent programs written
  in traditional imperative languages such as C/C++ and also in
  synchronous languages such as Esterel. It has been used as a
  transition step in translating imperative programs to hardware
  descriptions\cite{ahir-thesis}. The resulting high-level synthesis
  flow uses the labelling scheme to reuse hardware resources in the
  absence of timing information.

@@

A firing sequence is a sequence of transitions $\sigma = [t_{i_0},
t_{i_1}, \ldots]$ in $T$ that may fire in the petri net $N$ starting
at some marking $M$. $L(N)$ is the set of firing sequences possible in
the petri net $N$ starting from the initial marking $M_0$.

@@

Consider a simple place $p$ in a live and safe petri net $N =
(P,T,E,M)$. Let $u,v$ be the predecessor and successor transitions of
$p$ in $N$, i.e, $u,v \in T$ and $(u,p), (p,v)\in E$. For any firing
sequence $\sigma \in L(N)$, $u$ must occur before any occurence of
$v$, and occurences of $u$ and $v$ must strictly alternate in the
entire sequence.

The petri net $N' = N - \{p,u,v\}$ can be modelled as an initially
marked simple place $p_0$ that forms a cycle with $u$, $v$ and $p$.
The cycle $N'$ is live and safe. Every sequence in $L(N)$ can be
mapped to a sequence in $L(N')$ by simply deleting all transitions
other than $u$ and $v$, if and only if the petri net $N$ is live and
safe.

Similarly, relative to a simple transition $t$, the rest of the petri
net can be modelled as an initially marked simple place $p_0$ that
forms a simple cycle with $t$. Both replacements are valid if and only
if the original petri net is live and safe; the resulting simple cycle
is also live and safe. We use this simplification to examine the
effect of applying a construction rule on a Type-1 petri net:
